---
title: "tester"
author: "Kyle Wu"
date: "2023-02-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(AER)
library(dynlm)
library(forecast)
library(readxl)
library(stargazer)
library(scales)
library(quantmod)
library(urca)
```


### Detrended Total Vehicle Sales

```{r}
total_log <- ts(log(total_vehicles$TOTALNSA), start = c(1976,01), frequency = 12)

```

```{r}
total_log <- ts(log(total_vehicles$TOTALNSA), start = c(1976,01), frequency = 12)

diff_total_log <- diff(total_log)

decomp <- seas(total_log, x11 = "")

detrended <- total_log - decomp$data[, "trend"]

plot(total_log)
plot(detrended)
plot(diff_total_log)
adf.test(total_log)

kpss.test(diff_total_log, null = "Trend")

kpss.test(detrended, null = "Trend")
adf.test(diff_total_log)
attempt <- dynlm(detrended ~ L(detrended, 1) + L(detrended, 12))

model0 <- dynlm(diff_total_log ~ 1)
model11 <- dynlm(diff_total_log ~ season(diff_total_log), data = total_log)
anova(model0, model11)

plot(diff_total_log) +

lines(fitted(model0), col = 2) +

lines(fitted(model11), col = 4)
```



### Total Vehicles and Bank Prime Rate Loan:
```{r}
total

model2 <- dynardl()
```

```{r}
rate <- ts(bank_rate$MPRIME, start = c(1949, 1), frequency = 12)
kpss.test(rate, null = "Trend")

rate <- diff(rate)
kpss.test(rate, null = "Trend")
adf.test(rate)

total_rate_analysis <- ts.intersect(diff_total_log, rate)

plot(rate)

autoplot(total_rate_analysis)

model1 <- dynlm(diff_total_log ~ L(diff_total_log, 1:14) + L(rate, 0:15) + season(diff_total_log), data = total_rate_analysis)

summary(model1)

auto_ardl(diff_total_log ~ rate, data = total_rate_analysis, max_order = 20)

checkresiduals(model1)
```

```{r}
try_dat <- combined %>% filter(DATE > "1976-01-01") %>%
  mutate(diff_total_log = difference(difference(log(TOTALNSA, 12))),
         diff_rate = difference(MPRIME))

try_fit <- auto.arima(try_dat$diff_total_log, xreg = try_dat$diff_rate, stationary = TRUE)

checkresiduals(try_fit)
```

```{r}
stuff <- combined %>% filter(DATE >= "1976-01-01") %>%
  model(
    lag0 = ARIMA(TOTALNSA ~ pdq(d = 0) + MPRIME),
    lag1 = ARIMA(TOTALNSA ~ pdq(d = 0) + MPRIME + lag(MPRIME)),
    lag2 = ARIMA(TOTALNSA ~ pdq(d = 0) + MPRIME + lag(MPROME) + lag(MPRIME, 2))
  )

glance(stuff)
```


### Total Vehicles and New Car CPI:
```{r}
car_cpi <- ts(log(cpi_newcar$CUSR0000SETA01), start = c(1953, 1), frequency = 12)
kpss.test(car_cpi, null = "Trend")

car_cpi <- diff(car_cpi)
kpss.test(car_cpi, null = "Trend")
adf.test(car_cpi)

total_rate_analysis <- ts.intersect(diff_total_log, car_cpi)

plot(rate)

autoplot(total_rate_analysis)

model1 <- dynlm(diff_total_log ~ L(diff_total_log, 1:3) + L(rate, 0:5), data = total_rate_analysis)

summary(model1)
```

### Total Vehicles and Gas CPI:
```{r}
gas_cpi <- ts(log(fuel_prices$CUSR0000SETB01), start = c(1968, 2), frequency = 12) 
kpss.test(gas_cpi, null = "Trend")

gas_cpi <- diff(gas_cpi, 1)
kpss.test(gas_cpi, null = "Trend")
adf.test(gas_cpi)

total_gas_analysis <- ts.intersect(diff_total_log, gas_cpi)

autoplot(total_gas_analysis)

model6 <- dynlm(diff_total_log ~ L(diff_total_log, c(1:12)) + L(gas_cpi, 0:12) + season(diff_total_log), data = total_gas_analysis)

summary(model6)

checkresiduals(model6)

auto_ardl(diff_total_log ~ gas_cpi, data = total_gas_analysis, max_order = 24)
```

```{r}
gas_car_dat <- combined %>% filter(DATE >= "1976-01-01") %>%
  mutate(diff_total_log = difference(log(TOTALNSA),12) %>% difference(),
         diff_gas_cpi = difference(CUSR0000SETB01))

fit <- gas_car_dat %>%
  model(
    lag0 = ARIMA(diff_total_log ~ pdq(d = 0) + diff_gas_cpi),
    lag1 = ARIMA(diff_total_log ~ pdq(d = 0) + diff_gas_cpi + lag(diff_gas_cpi)),
    lag2 = ARIMA(diff_total_log ~ pdq(d = 0) + diff_gas_cpi + lag(diff_gas_cpi) + lag(diff_gas_cpi, 2)),
    lag3 = ARIMA(diff_total_log ~ pdq(d = 0) + diff_gas_cpi + lag(diff_gas_cpi) + lag(diff_gas_cpi, 2) + lag(diff_gas_cpi, 3)))

glance(fit)

fit_best <- gas_car_dat %>%
  model(ARIMA(diff_total_log ~ pdq(d = 0) + diff_gas_cpi + lag(diff_gas_cpi) + lag(diff_gas_cpi, 2) + lag(diff_gas_cpi, 3)))

fit_best %>% gg_tsresiduals()
report(fit_best)
augment(fuel_fit) %>% filter(.model == 'search') %>%
  features(.innov, ljung_box, lag = 10, dof = 3)

augment(fit_best) %>% features(.innov, ljung_box, lag = 10, dof = 3)
```
### Total Vehicle and Unemployment Rate:
```{r}
unemp <- ts(unemployment$UNRATE, start = c(1948, 1), frequency = 12)

unemp <- diff(unemp)

kpss.test(unemp, null = "Trend")
adf.test(unemp)

autoplot(unemp)

total_unemp_analysis <- ts.intersect(diff_total_log, unemp)

autoplot(total_unemp_analysis)

model8 <- dynlm(diff_total_log ~ L(diff_total_log, 1:12) + L(unemp, 0:1) + season(diff_total_log), data = total_unemp_analysis)

summary(model8)

checkresiduals(model8)
```
```{r}
(fit <- auto.arima(uschange[,"Consumption"],
  xreg=uschange[,"Income"]))

veh_unemp_fit <- auto.arima(model8, xreg = model.matrix(model8))
```


### Total Vehicles and Domestic Auto Production:
```{r}

auto_prod <- ts(log(auto_production$DAUPNSA), start = c(1993, 1), frequency = 12)

kpss.test(auto_prod, null = "Trend")

auto_prod <- diff(auto_prod, 1)
kpss.test(auto_prod, null = "Trend")
adf.test(auto_prod)

total_prod_analysis <- ts.intersect(diff_total_log, auto_prod) 

model51 <- dynlm(diff_total_log ~ L(diff_total_log, 1:10) + L(auto_prod, 0:10) + season(total_log), data = total_prod_analysis)

summary(model51)

Anova(model5, model51)

plot(model51)
autoplot(total_prod_analysis)

checkresiduals(model51)


auto_ardl(diff_total_log ~ auto_prod, data = total_prod_analysis, max_order = 24)

order <- 1:24
sapply
BICs <- base::sapply(order, function(x)
  BIC(dynlm(diff_total_log ~ L(diff_total_log, 1:x) + L(auto_prod, 1:x))))

```

```{r}

xreg <- model.matrix(model51, newdata = total_prod_analysis)

model_arima <- auto.arima(total_prod_analysis[, "diff_total_log"], xreg = xreg)

plot(total_prod_analysis[, "diff_total_log"])
```


```{r}
mod_data <- combined %>% filter(DATE >= "1976-01-01")
mod <- lm(TOTALNSA ~ MPRIME, data = mod_data)
```

```{r}
auto_production %>%
  mutate(diff_prod = difference(production_adjusted, 12) %>% difference()) %>%
  features(diff_prod, unitroot_kpss) %>%
  gt::gt()

diff1 <- autoplot(ts(difference(auto_production$production_adjusted), start = c(1993,01), frequency = 12))
diff_season <- autoplot(ts(difference(auto_production$production_adjusted, 12) %>% difference(), start = c(1993,01), frequency = 12))

grid.arrange(diff1, diff_season)
```




After we have successfully differenced our data, we can then begin to
consider what types of models to fit to our stationary time series. When
considering ARIMA models, there are two model types that we can use:
autoregressive models and moving average models.





We will first discuss autoregressive models. In autoregression, our goal
is to forecast the variable of interest using linear combinations of
previous values of the variable. This means that in an autoregression,
the variable is regressed on itself. An model that utilizes
autoregression of order p can be denoted by
$$y_t = c + \phi_1y_{t-1} + \phi_2y_{t-2} + \cdots + \phi_py_{t-p} + \epsilon_t$$,
where $\epsilon_t$ denotes white noise. In such a case, we would say
that our model is an AR(p) model. For these types of models, we face
certain constraints, mainly:

-   For an AR(1) model: $-1 < \phi_1 < 1$

-   For an AR(2) model:
    $-1 < \phi_2 < 1, \phi_1 + \phi_2 < 1, \phi_2-\phi_1 < 1$.

While autoregression models use past values of itself, moving average
models use past forecast errors in a regression_like model
[@forecastbook]. Moving average models take on the form
$$y_t = c + \epsilon_t + \theta_1\epsilon_{t-1} + \theta_2\epsilon_{t-2} + \cdots + \theta_q\epsilon_{t-q}$$,
where $\epsilon_t$ is white noise. In such a case, we would say that our
model is an MA(q) model. Each value of $y_t$ can be thought of as a
weighted moving average of past forecast errors. MA models also have
stationarity constraints which are as follows:

-   For an MA(1) model: $-1 < \theta_1 < 1$

-   For an MA(2) model:
    $-1 < \theta_2 < 1, \theta_1 + \theta_2 >1, \theta_1-\theta_2 < 1$.
    
    
### Non-Seasonal ARIMA Models

For our non-seasonal time series we can appropriately use Non-seasonal
ARIMA models which follow the form
$$y_t' = c + \phi_1y_{t-1}' + \cdots + \phi_py_{t-p}' + \theta_1\epsilon_{t-1} + \cdots + \theta_1\epsilon_{t-q} + \epsilon_t$$

To find the best model, we will have to create ACF and PACF plots so
that we may visually decide which ARIMA model may best model our
stationary time series, we will also take advantage of functions which
will pick the best ARIMA model and then compare the given answers. The
ACF will allow us to decide which MA nmodel to use and the PACF will
allow us to decide which AR model to use.

In the case of fuel prices:

```{r}
fuel_prices %>%
  gg_tsdisplay(difference(fuel_adjusted), plot_type = 'partial')

fit <- fuel_prices %>%
  model(ARIMA(fuel_adjusted))

report(fit)
```

```{r}
fuel_fit <- fuel_prices %>%
  model(arima310 = fable::ARIMA(fuel_adjusted ~ pdq(2,1,0)),
        arima011 = fable::ARIMA(fuel_adjusted ~ pdq(0,1,1)),
        stepwise = fable::ARIMA(fuel_adjusted),
        search = fable::ARIMA(fuel_adjusted, stepwise = FALSE))
fuel_fit

glance(fuel_fit) %>% arrange(AICc) %>% select(.model:BIC)

fuel_fit %>% select(search) %>% gg_tsresiduals()

augment(fuel_fit) %>% filter(.model == 'search') %>%
  features(.innov, ljung_box, lag = 10, dof = 3)
```

### Seasonal ARIMA Models

Combining autoregression and moving average models along with
differencing allows us to obtain ARIMA models, which stands for Auto
Regressive Integrated Moving Averge. ARIMA models can be written as

```{r}
auto_production_ts


auto_production %>%
  gg_tsdisplay(difference(production_adjusted, 12) %>% difference(), plot_type = 'partial', lag = 36)       

auto_production <- auto_production %>% mutate(production_log_adjust =  difference(log(DAUPNSA), 12) %>% difference())

fit3 <- auto_production %>% filt
  model(ARIMA())

fit3 <- auto.arima(diff(log(auto_production$DAUPNSA), 12) %>% diff())

fit_things <- auto_production %>% model(ARIMA(production_log_adjust))

report(fit_things)

gg_tsresiduals(fit_things)

augment(fit_things) %>% features(.innov, ljung_box, lag = 10, dof = 3)
forecast(fit3, h=10)

fit_things %>% forecast(h = 24) %>% autoplot(auto_production)

autoplot(auto_production)
plot(forecast(fit3, h=30))

trend <- auto

thingy <- auto.arima(auto_production$DAUPNSA)

auto_production2 <- auto_production %>% filter(DATE <= "2020-01-01")

thingy2 <-auto.arima(auto_production2$DAUPNSA)

auto_production2 <- auto_production2 %>%
  mutate(pre_difference = difference(auto_prodction2$DAUPNSA))
  
plot(auto_prodction2$pre_difference)

thingy.forecast <- forecast(thingy, level = c(95), h = 50)
thingy2.forecast <- forecast(thingy2, level = c(95), h = 36)

autoplot(thingy.forecast)
autoplot(thingy2.forecast)

grid.arrange(autoplot(auto_production), autoplot(thingy2.forecast))

a <- forecast(fit3, level = c(95), h = 50)
autoplot(a + seasonaldummy(auto_production))
```

```{r, include=FALSE}
fit <- auto.arima(diff(fuel_prices$CUSR0000SETB01))
fit
forecasting <- forecast(fit, h = 30) 

plot(forecasting)

```

```{r}
total_vehicles %>%
  ACF(difference(total_adjusted, 12) %>% difference()) %>%
  autoplot() + labs(title = "Total Vehicles Sold")

total_vehicles %>%
  mutate(diff_total = difference(total_adjusted)) %>%
  features(diff_total, unitroot_kpss)

total_vehicles %>%
  features(total_adjusted, unitroot_ndiffs)

total_vehicles %>% model(ARIMA(total_adjusted)) %>% report()
```

After we have appropriately differenced our models so that they are
stationary, we can then begin to consider what type of models to fit to
our data.


```{r, include=FALSE}
total_vehicles$TOTALNSA <- ts(total_vehicles$TOTALNSA, start = c(1976,01), frequency = 12)

plot(total_vehicles$TOTALNSA)

dopen <- diff(total_vehicles$TOTALNSA, 1)

ddopen <- diff(dopen, 1)

plot(dopen)
plot(ddopen)
ar1_model<- auto.arima(ddopen)

ar1_model

forecasts_AR1<- forecast(ar1_model, h = 30)

plot(forecasts_AR1)
```

```{r, include=FALSE}
cpi_newcar$CUSR0000SETA01 <- ts(cpi_newcar$CUSR0000SETA01, start = c(1953,01), frequency = 12)

plot(cpi_newcar$CUSR0000SETA01)

dopen2 <- diff(cpi_newcar$CUSR0000SETA01, 1)

ddopen2 <- diff(dopen2, 1)

plot(dopen2)
plot(ddopen2)
ar2_model<- auto.arima(dopen2)

ar2_model

forecasts_AR2<- forecast(ar2_model, h = 30)

plot(forecasts_AR2)
```

```{r, include=FALSE}
plot(auto_production$DAUPNSA)

tre <- time(auto_production$DAUPNSA) - mean(time(auto_production$DAUPNSA))

tre2 <- tre^2

regmodel2 = lm(auto_production$DAUPNSA ~ tre)

summary(regmodel2)

acf2(resid(regmodel2))

adjreg2 = sarima(auto_production$DAUPNSA, 3, 1, 3, xreg = cbind(tre))

adjreg2$fit$coef


```

```{r, include=TRUE}
decomp_tv <- total_vehicles %>%
  model(stl = STL(TOTALNSA))

components(decomp_tv) %>% autoplot()
```

```{r, include=FALSE}
trial <- autoplot(combined %>% filter(DATE > "1975-12-01"), TOTALNSA)
```

```{r, message=FALSE, warning=FALSE}
GGally::ggpairs(combined[,2:6])
```

```{r, include=FALSE}
require(urca)
total_vehicles$TOTALNSA %>% ur.kpss %>% summary() 
auto_production$DAUPNSA %>% ur.kpss %>% summary()
bank_rate$MPRIME %>% ur.kpss %>% summary()
fuel_prices$CUSR0000SETB01 %>% ur.kpss %>% summary()
```



```{r, include=FALSE}
plot(ts(auto_production$production_adjusted))

trend = time(auto_production$production_adjusted) - mean(time(auto_production$production_adjusted))

regmodel = lm(auto_production$production_adjusted ~ trend)

summary(regmodel)

acf2(resid(regmodel))

adjreg = sarima(auto_production$production_adjusted, 0,1,3, xreg = cbind(trend))

```

```{r, include=FALSE}
total_vehicles %>% features(log(TOTALNSA), unitroot_kpss)
auto_production %>% features(log(DAUPNSA), unitroot_kpss)
bank_rate %>% features(MPRIME, unitroot_kpss)
fuel_prices %>% features(log(CUSR0000SETB01), unitroot_kpss)
unemployment %>% features(UNRATE, unitroot_kpss)
unemployment %>% features(UNRATE, unitroot_kpss)
```

```{r}
# Total Auto Production and Bank Prime Rate Loans from 1976-01-01
tap_bpm_data <- combined %>% filter(DATE >= "1976-01-01")
tap_bpm <- dynlm::dynlm(TOTALNSA ~ MPRIME + L(TOTALNSA), data = tap_bpm_data)
summary(tap_bpm)

plot.ts(tap_bpm_data$TOTALNSA)
lines(fitted(tap_bpm), col = "red")

trys <- dynlm(total_vehicles$TOTALNSA ~ Lag(total_vehicles$TOTALNSA, 1) + Lag(total_vehicles$TOTALNSA, 12))

plot.ts(total_vehicles$TOTALNSA)
lines(fitted(trys), col = "red")
```

```{r}
tap_bpm <- combined %>%
  filter(DATE >= "1976-01-01") %>% 
  model(ARIMA(TOTALNSA ~ MPRIME + lag(MPRIME) + lag(MPRIME, 2) + lag(MPRIME, 3)))

report(tap_bpm)

tap_bpm %>% gg_tsresiduals()
```

```{r}
augment(tap_bpm) %>%
  features(.innov, ljung_box, dof = 3, lag = 24)
```


### Trying to see what code does

```{r}
USMacroSWQ <- read_xlsx("Data/us_macro_quarterly.xlsx",
                         sheet = 1,
                         col_types = c("text", rep("numeric", 9)))

# format date column
USMacroSWQ$X__1 <- as.yearqtr(USMacroSWQ$X__1, format = "%Y:0%q")

# adjust column names
colnames(USMacroSWQ) <- c("Date", "GDPC96", "JAPAN_IP", "PCECTPI", 
                          "GS10", "GS1", "TB3MS", "UNRATE", "EXUSUK", "CPIAUCSL")
```

```{r}
data("USDistLag", package = "lmtest")
dfm1 <- dynlm(consumption ~ gnp + L(consumption), data = USDistLag)
dfm2 <- dynlm(consumption ~ gnp + L(gnp), data = USDistLag)
plot(USDistLag[, "consumption"]) +
lines(fitted(dfm1), col = 2) +
lines(fitted(dfm2), col = 4)
if(require("lmtest")) encomptest(dfm1, dfm2)
```

